---
title: "Valentine's Day Spending Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Valentine's Day Spending Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```

## Introduction of `Valentindtrend` package

This package includes four datasets: `gifts_age`, `gifts_gender`, `historical_full_spending`, and `historical_gifts_trends_percent_buying`. These datasets show annual trends in how people plan to celebrate Valentine’s Day over the past decade, including total expected spending, average per-person spending, and preferred gift types, broken down by age and gender."

The data can be accessed by loading the `Valentinetrend` and calling on the object `gifts_age`, or alternatively using `Valentinetrend::gifts_age` if you want to analyse Valentine's Day spending trends by different age group. 

## Usage
```{r, eval = FALSE}
library(Valentinetrend)

# Load the gifts_age dataset
head(gifts_age)

# Alternatively, use the full namespace
head(Valentinetrend::gifts_age)
```


```{r setup}
library(Valentinetrend)
library(knitr)
library(ggplot2)
library(dplyr)
library(RColorBrewer)
library(tidyr)
```

## Examples

Some examples analyses using the datasets are provided. You'll learn how to perform basic data exploration and create visualizations by using this package!

Below, we use the `historical_full_spending` dataset to create a combined bar and line chart showing total expected and average per person spending from 2010 to 2022.

```{r}
ggplot(historical_full_spending, aes(x = Year)) +
      geom_bar(
        aes(y = PerPerson, text = paste0("Year: ", Year, "<br>Per Person Spend: $", round(PerPerson, 2))),
        stat = "identity", fill = "hotpink", alpha = 0.7
      ) +
      geom_line(
        aes(y = total_expected_valentines_day_spending_in_billions * 10),
        color = "mediumvioletred", size = 1
      ) +
      geom_point(
        aes(y = total_expected_valentines_day_spending_in_billions * 10,
            text = paste0("Year: ", Year, "<br>Total Spend: $", total_expected_valentines_day_spending_in_billions, "B")),
        color = "mediumvioletred", size = 2
      ) +
      geom_text(
        aes(y = total_expected_valentines_day_spending_in_billions * 10 + 10,
            label = paste0("$", total_expected_valentines_day_spending_in_billions, "B")),
        size = 3, fontface = "bold"
      ) +
      labs(
        title = "Valentine's Day Spending Plans",
        x = "Year",
        y = "Spending ($)"
      ) +
      theme_minimal() +
      theme(plot.title = element_text(hjust = 0.5, size = 16)) +
  annotate("point", x = 2013.9, y = -12, color = "hotpink", size = 3) +  
  annotate("text", x = 2015.1, y = -12, label = "Per Person Spend", size = 3, color = "black") + 
  annotate("point", x = 2017, y = -12, color = "mediumvioletred", size = 3) + 
  annotate("text", x = 2017.8, y = -12, label = "Total Spend", size = 3, color = "black")

```

We also create a separate section for each dataset, showing different analyses or visualizations.

### Exploring the spending trends on gifts breaking down by age

The `gifts_age` dataset provides insights into spending trends by different age groups. Here, we create a line chart to visualize gift preferences among different age groups.

```{r}
# Use pivot_longer to reshape the data
gifts_age_long <- gifts_age %>%
  pivot_longer(cols = Candy:GiftCards, 
               names_to = "GiftCategory", 
               values_to = "Percentage")

# Create the line plot for gift categories
ggplot(gifts_age_long, aes(x = Age, y = Percentage, color = GiftCategory, group = GiftCategory)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  labs(title = "Spending on Different Gifts By Age Group",
       x = "Age Group", 
       y = "Percentage (%)",
       color = "Gift Category") +
  theme_minimal()+
  scale_color_brewer(palette = "Accent")
```

## Interactive plots

Alternatively, you can create interactive chart using the `plotly` package, as demonstrated in the Shiny app. You can launch the app as shown in below to explore engaging interactive plots that provide a more dynamic and user-friendly experience!

```{r, eval = FALSE}
Valentinetrend::launch_app()
```


For example, to add **plotly** functionality with hover info to the chart, use the following code:

```{r, eval = FALSE}
# Copy the code from the 'historical_full_spending' combined chart
# Example:
# p <- ggplot(historical_full_spending, aes(x = Year)) + 
#   ... +
#   theme(plot.title = element_text(hjust = 0.5, size = 16))

# To make the chart interactive, add the following code:
ggplotly(p, tooltip = "text") %>%
  layout(
    hovermode = "closest"
  ) %>%
  style(hoverinfo = "none", traces = 4)
``` 


The above guide demonstrates how to use the `Valentinetrend` package to analyze Valentine's Day spending trends. You can also explore additional datasets (`gifts_gender`, `historical_gifts_trends_percent_buying`) for further insights. 
