---
title: "Valentine's Day Consumer Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Valentine's Day Consumer Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```

This package includes four datasets: `gifts_age`, `gifts_gender`, `historical_full_spending`, and `historical_gifts_trends_percent_buying`. These datasets show annual trends in how people plan to celebrate Valentine’s Day over the past decade, including total expected spending, average per-person spending, and preferred gift types, broken down by age and gender."

The data can be accessed by loading the `Valentinetrend` and calling on the object `gifts_age`, or alternatively using `Valentinetrend::gifts_age` if you want to analyse Valentine's Day spending trends by different age group. 

```{r}
library(Valentinetrend)

# Load the gifts_age dataset
head(gifts_age)

# Alternatively, use the full namespace
head(Valentinetrend::gifts_age)
```


```{r setup}
library(Valentinetrend)
library(knitr)
library(ggplot2)
library(dplyr)
library(RColorBrewer)
library(tidyr)
```

# Example

Some examples analyses using the datasets are provided. You'll learn how to perform basic data exploration and create visualizations by using this package!

Below, we use the `historical_full_spending` dataset to create a combined bar and line chart showing total expected and average per person spending from 2010 to 2022.

```{r}
ggplot(historical_full_spending, aes(x = Year)) +
  geom_bar(aes(y = PerPerson), stat = "identity", fill = "hotpink") +
  geom_line(aes(y = total_expected_valentines_day_spending_in_billions * 10), color = "mediumvioletred", size = 1) +  
  geom_point(aes(y = total_expected_valentines_day_spending_in_billions * 10), color = "mediumvioletred", size = 2) +
  geom_text(aes(y = total_expected_valentines_day_spending_in_billions * 10, 
                label = paste0("$", total_expected_valentines_day_spending_in_billions, "B")),
            vjust = -0.4, color = "black", size = 3.6)+  
  labs(title = "Valentine's Day Spending Trends",
       y = "Spending ($)") +
  scale_fill_manual(values = c("Per Person Spend" = "hotpink")) +  # Specify fill color for bars
  scale_color_manual(values = c("Total Spend" = "mediumvioletred")) + 
  scale_x_continuous(breaks = historical_full_spending$Year) +
  theme_classic()+
  theme(plot.title = element_text(hjust = 0.5, size = 16),
        legend.position = "none")+
  annotate("point", x = 2014, y = -12, color = "hotpink", size = 3) +  
  annotate("text", x = 2015.1, y = -12, label = "Per Person Spend", size = 3, color = "black") + 
  annotate("point", x = 2017, y = -12, color = "mediumvioletred", size = 3) + 
  annotate("text", x = 2017.8, y = -12, label = "Total Spend", size = 3, color = "black") 

```

Create a separate section for each dataset, showing different analyses or visualizations.

## Exploring the spending trends on gifts breaking down by age

The `gifts_age` dataset provides insights into spending trends by different age groups. Here, we create a line chart to visualize gift preferences among different age groups.

```{r}
# Use pivot_longer to reshape the data
gifts_age_long <- gifts_age %>%
  pivot_longer(cols = Candy:GiftCards, 
               names_to = "GiftCategory", 
               values_to = "Percentage")

# Create the line plot for gift categories
ggplot(gifts_age_long, aes(x = Age, y = Percentage, color = GiftCategory, group = GiftCategory)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  labs(title = "Spending on Different Gifts By Age Group",
       x = "Age Group", 
       y = "Percentage (%)",
       color = "Gift Category") +
  theme_minimal()+
  scale_color_brewer(palette = "Accent")
````


Alternatively, you can create interactive charts using the plotly package, as demonstrated in the Shiny app. You can launch the app with `Valentinetrend::launch_app()` to explore engaging interactive plots.

The above guide demonstrates how to use the `Valentinetrend` package to analyze Valentine's Day spending trends. You can also explore additional datasets (`gifts_gender`, `historical_gifts_trends_percent_buying`) for further insights. For more detailed documentation, visit the [pkgdown site](https://your-username.github.io/Valentinetrend).
