---
title: "Valentine's Day Spending Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Valentine's Day Spending Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```

## Introduction of `Valentindtrend` package
<br>
This package includes four datasets: `gifts_age`, `gifts_gender`, `historical_full_spending`, and `historical_gifts_trends_percent_buying`. These datasets show annual trends in how people plan to celebrate Valentine’s Day over the past decade, including total expected spending, average per-person spending, and preferred gift types, broken down by age and gender."

The data can be accessed by loading the `Valentinetrend` and calling on the object `gifts_age`, or alternatively using `Valentinetrend::gifts_age` if you want to analyse Valentine's Day spending trends by different age group. 

## Usage
```{r, eval = FALSE}
library(Valentinetrend)

# Load the gifts_age dataset
head(gifts_age)

# Alternatively, use the full namespace
head(Valentinetrend::gifts_age)
```


```{r setup}
library(Valentinetrend)
library(knitr)
library(ggplot2)
library(dplyr)
library(RColorBrewer)
library(tidyr)
```

## Examples

Some examples analyses using the datasets are provided. You'll learn how to perform basic data exploration and create visualizations by using this package!

Below, we use the `historical_full_spending` dataset to create a combined bar and line chart showing total expected and average per person spending from 2010 to 2022.

```{r}
# Create a bar and line chart to visualize Valentine's Day spending trends over the years
ggplot(historical_full_spending, aes(x = Year)) +
  
  # Add a bar chart to show the average per-person spending each year
  geom_bar(
    aes(y = PerPerson, 
        text = paste0("Year: ", Year, "<br>Per Person Spend: $", round(PerPerson, 2))),  # Add hover text for plotly
    stat = "identity",  # Use identity to represent actual values
    fill = "hotpink",   # Set bar color to hotpink
    alpha = 0.7         # Set transparency for the bars
  ) +
  
  # Add a line to represent total expected spending (scaled by 10 for visualization)
  geom_line(
    aes(y = total_expected_valentines_day_spending_in_billions * 10),  # Scale total spending for better visualization
    color = "mediumvioletred",  # Line color set to mediumvioletred
    size = 1                    # Set line thickness
  ) +
  
  # Add points for each total expected spending data point
  geom_point(
    aes(y = total_expected_valentines_day_spending_in_billions * 10,  # Match y value to scaled total spending
        text = paste0("Year: ", Year, "<br>Total Spend: $", total_expected_valentines_day_spending_in_billions, "B")),  # Add hover text for plotly
    color = "mediumvioletred",  # Set point color to match the line
    size = 2                    # Set point size
  ) +
  
  # Add text labels above the total spending points for additional information
  geom_text(
    aes(y = total_expected_valentines_day_spending_in_billions * 10 + 10,  # Adjust y position for text labels
        label = paste0("$", total_expected_valentines_day_spending_in_billions, "B")),  # Show total spending in billions
    size = 3,          # Set text size
    fontface = "bold"  # Make text bold
  ) +
  
  # Add plot titles and axis labels
  labs(
    title = "Valentine's Day Spending Plans",  # Set plot title
    x = "Year",                                # Set x-axis label
    y = "Spending ($)"                         # Set y-axis label
  ) +
  
  # Apply a minimal theme for a clean layout
  theme_minimal() +
  
  # Center the plot title and increase font size
  theme(plot.title = element_text(hjust = 0.5, size = 16)) +
  
  # Add manual annotations to represent legend information
  annotate("point", x = 2013.9, y = -12, color = "hotpink", size = 3) +  # Add point for "Per Person Spend"
  annotate("text", x = 2015.1, y = -12, label = "Per Person Spend", size = 3, color = "black") +  # Add label for "Per Person Spend"
  annotate("point", x = 2017, y = -12, color = "mediumvioletred", size = 3) +  # Add point for "Total Spend"
  annotate("text", x = 2017.8, y = -12, label = "Total Spend", size = 3, color = "black")  # Add label for "Total Spend"

```

We also create a separate section for each dataset, showing different analyses or visualizations.

### Spending trends on gifts breaking down by age

The `gifts_age` dataset provides insights into spending trends by different age groups. Here, we create a line chart to visualize gift preferences among different age groups.

```{r}
# Reshape the 'gifts_age' data from wide to long format using pivot_longer
# This converts columns 'Candy' to 'GiftCards' into a single 'GiftCategory' column
# and their respective values into a 'Percentage' column
gifts_age_long <- gifts_age %>%
  pivot_longer(
    cols = Candy:GiftCards,          # Specify columns to reshape
    names_to = "GiftCategory",       # New column for gift categories
    values_to = "Percentage"         # New column for spending percentages
  )

# Create a line plot to visualize spending on different gift categories by age group
ggplot(gifts_age_long, aes(x = Age, y = Percentage, color = GiftCategory, group = GiftCategory)) +
  
  # Add a line for each gift category, grouped by age group
  geom_line(size = 1) +
  
  # Add points to indicate data points for each age group and gift category
  geom_point(size = 2) +
  
  # Add titles and labels to the plot
  labs(
    title = "Spending on Different Gifts By Age Group",  # Plot title
    x = "Age Group",                                     # X-axis label
    y = "Percentage (%)",                                # Y-axis label
    color = "Gift Category"                              # Legend title
  ) +
  
  # Apply minimal theme for a clean look
  theme_minimal() +
  
  # Use a color palette for distinguishing different gift categories
  scale_color_brewer(palette = "Accent")

```

## Interactive plots

Alternatively, you can create interactive chart using the `plotly` package, as demonstrated in the Shiny app. You can launch the app as shown in below to explore engaging interactive plots that provide a more dynamic and user-friendly experience!

```{r, eval = FALSE}
Valentinetrend::launch_app()
```


For example, to add **plotly** functionality with hover info to the chart, use the following code:

```{r, eval = FALSE}
# Copy the code from the 'historical_full_spending' combined chart
# Example:
# p <- ggplot(historical_full_spending, aes(x = Year)) + 
#   ... +
#   theme(plot.title = element_text(hjust = 0.5, size = 16))

# To make the chart interactive, add the following code:
ggplotly(p, tooltip = "text") %>%
  layout(
    hovermode = "closest"
  ) %>%
  style(hoverinfo = "none", traces = 4)
``` 


The above guide demonstrates how to use the `Valentinetrend` package to analyze Valentine's Day spending trends. You can also explore additional datasets (`gifts_gender`, `historical_gifts_trends_percent_buying`) for further insights. 
